---
title: "p8105_hw6_sw3879"
author: "Siqing Wang"
date: "2023-11-30"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(tidyverse)
library(ggplot2)
```

## Problem 1
Reading and cleaning dataset
```{r}
homicide_df = read_csv("data/homicide-data.csv") |> 
  janitor::clean_names() |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
  ) |> 
   filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) |> 
   filter(victim_race %in% c("White", "Black"))
```

using `glm` for logistic regression for Baltimore
```{r}
baltimore_glm = homicide_df |> filter(city_state == "Baltimore, MD") |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 

baltimore_glm |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate),
         upperCI = exp(estimate + qnorm(.975)*std.error),
         lowerCI = exp(estimate + qnorm(.025)*std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(term, OR, upperCI, lowerCI) |> 
    knitr::kable(digits = 3)
```

Interpretation: In Baltimore MD, male victims have 0.426 times the odds to have resolved case compared with female victims while all other variables held fixed.  

Using `glm` for all cities
```{r}
nest_glm = homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    glm_model = map(data, \(df) glm(resolved ~ victim_age + victim_race + victim_sex, data = df)),
    glm_result = map(glm_model, broom::tidy)
  ) |> 
  select(-data, -glm_model) |> 
  unnest(glm_result) |> 
  filter(term == "victim_sexMale") |> 
  mutate(OR = exp(estimate),
         upperCI = exp(estimate + qnorm(.975)*std.error),
         lowerCI = exp(estimate + qnorm(.025)*std.error),
         city_state = fct_reorder(city_state, OR)) |>
  select(city_state, term, OR, upperCI, lowerCI)
```

Creating a plot
```{r}
nest_glm |> ggplot(aes(x = city_state, y = OR)) +
  geom_point() + geom_errorbar(aes(ymin = lowerCI, ymax = upperCI)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Estimated OR for resolving homicides comparing male to female victims")
```

NYC has the lowest odd ratio while Albuquerque has the highest, meaning that in NYC, the odds of resolved cases for male victims vs. female victims is the lowest, and in Albuquerque the highest, while all other factors held constant. 

## Problem 2
Reading the data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Setting up bootstrap function
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```

Draw 5000 samples
```{r}
set.seed(1)

boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df))
  )
```

Do `lm` fit
```{r}
boot_results = 
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_tidy = map(models, broom::tidy),
    results_glance = map(models, broom::glance)
  ) |> 
  select(strap_number, results_tidy, results_glance) |> 
  unnest(results_tidy, results_glance)
```

```{r}
results_logb1b2 = boot_results |> 
  select(strap_number, term, estimate, std.error) |> 
  filter(term %in% c("tmin","prcp")) |> 
  pivot_wider(
     names_from = term,
     values_from = c(estimate, std.error)
  ) |> 
  mutate(
    logb1b2 = log(estimate_tmin * estimate_prcp)
  ) 

results_rsq = boot_results |> 
  select(strap_number, r.squared) |> 
  distinct(.keep_all = TRUE)
```

Plotting estimates
```{r}
results_logb1b2 |> ggplot(aes(y = logb1b2)) + geom_boxplot() +
  labs(title = "Estimation of log(b1*b2)")
```

The 95% CI for log(b1*b2)
```{r}
results_logb1b2 |> summarise(
  upperCI = quantile(logb1b2, 0.975, na.rm = TRUE),
  lowerCI = quantile(logb1b2, 0.025, na.rm = TRUE)) |> 
  knitr::kable()
```

```{r}
results_rsq |> ggplot(aes(y = r.squared)) + geom_boxplot() + 
  labs(title = "Esimation of R squared")
```

The 95% CI for r-squared
```{r}
results_rsq |> summarise(
  upperCI = quantile(r.squared, 0.975, na.rm = TRUE),
  lowerCI = quantile(r.squared, 0.025, na.rm = TRUE)) |> 
  knitr::kable()
```
The estimated R squared value is close to 1, meaning there is a high percentage in the variance of `tmax` that can be explained by the relationship with the predictors `tmin` and `prcp`. 

COMMENT
Significant amount of NaN values in log(b1*b2) due to negative precipitation values. 

## Problem 3

Reading and cleaning data
```{r}
bwt_df = read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("male", "female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
  mrace = factor(mrace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
  malform = factor(malform, levels = c(0, 1), labels = c("absent", "present"))
  )
```

My hypothesis is that birth weight is associated with presence of malformation `malform` and mother’s weight gain during pregnancy `wtgain`, since I hypothesize pregnant mom with greater weight gain may have babies that weigh more, and those who don't have malformation might have higher weight. I will use MLR to fit `malform` and `wtgain` as predictor, and `bwt` as outcome. I will also include an interaction term. 

```{r}
q3m1 = lm(bwt ~ malform + wtgain + malform*wtgain, data = bwt_df)

broom::tidy(q3m1)
```

Plotting residuals agains fitted value for my model:
```{r}
bwt_df |> modelr::add_predictions(q3m1) |> 
  modelr::add_residuals(q3m1) |> 
  ggplot(aes(x = pred, y = resid)) + geom_point() +
  labs(title = "MLR of birth weight with malformations and weight gain")
```

Now comparing with the other two given models:
```{r}
q3model2 = lm(bwt ~ blength + gaweeks, data = bwt_df)
broom::tidy(q3model2)

q3model3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = bwt_df)
broom::tidy(q3model3)
```

```{r}
cv_df = 
  bwt_df |> 
  modelr::crossv_mc(n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_results = cv_df |> mutate(
  q3m1 = map(train, \(df) lm(bwt ~ malform + wtgain + malform*wtgain, data = df)),
  q3m2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
  q3m3 = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = df))
  )|> 
    mutate(
    rmse_q3m1 = map2_dbl(q3m1, test, \(mod, df) modelr::rmse(mod, df)),
    rmse_q3m2 = map2_dbl(q3m2, test, \(mod, df) modelr::rmse(mod, df)),
    rmse_q3m3 = map2_dbl(q3m3, test, \(mod, df) modelr::rmse(mod, df))
    )

cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model_type",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  group_by(model_type) |> 
  summarize(m_rmse = mean(rmse))
```

Plot the RMSE for all three models
```{r}
cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model_type",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x = model_type, y = rmse)) +
  geom_violin() +
  labs(title = "Comparison between my model (q3m1) and 2 given models")
```

The 2 alternative models are significantly better than the model I proposed due to the lower RMSE. 
