---
title: "p8105_hw6_sw3879"
author: "Siqing Wang"
date: "2023-11-30"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(tidyverse)
library(ggplot2)
```

## Problem 1
Reading and cleaning dataset
```{r}
homicide_df = read_csv("data/homicide-data.csv") |> 
  janitor::clean_names() |> 
  mutate(
    city_state = paste(city, state, sep = ", "),
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
  ) |> 
   filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) |> 
   filter(victim_race %in% c("White", "Black"))
```

using `glm` for logistic regression for Baltimore
```{r}
baltimore_glm = homicide_df |> filter(city_state == "Baltimore, MD") |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 

baltimore_glm |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate),
         upperCI = exp(estimate + qnorm(.975)*std.error),
         lowerCI = exp(estimate + qnorm(.025)*std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(term, OR, upperCI, lowerCI) |> 
    knitr::kable(digits = 3)
```

Interpretation: In Baltimore MD, male victims have 0.426 times the odds to have resolved case compared with female victims while all other variables held fixed.  

Using `glm` for all cities
```{r}
nest_glm = homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    glm_model = map(data, \(df) glm(resolved ~ victim_age + victim_race + victim_sex, data = df)),
    glm_result = map(glm_model, broom::tidy)
  ) |> 
  select(-data, -glm_model) |> 
  unnest(glm_result) |> 
  filter(term == "victim_sexMale") |> 
  mutate(OR = exp(estimate),
         upperCI = exp(estimate + qnorm(.975)*std.error),
         lowerCI = exp(estimate + qnorm(.025)*std.error),
         city_state = fct_reorder(city_state, OR)) |>
  select(city_state, term, OR, upperCI, lowerCI)
```

Creating a plot
```{r}
nest_glm |> ggplot(aes(x = city_state, y = OR)) +
  geom_point() + geom_errorbar(aes(ymin = lowerCI, ymax = upperCI)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Estimated OR for resolving homicides comparing male to female victims")
```

NYC has the lowest odds ratio while Albuquerque has the highest, meaning that in NYC, the odds of resolved cases for male victims vs. female victims is the lowest, and in Albuquerque the highest, while all other factors held constant. 

## Problem 2
Reading the data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Setting up bootstrap function
```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```

Draw 5000 samples
```{r}
set.seed(1)

boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df))
  )
```

Do `lm` fit
```{r}
boot_results = 
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_tidy = map(models, broom::tidy),
    results_glance = map(models, broom::glance)
  ) |> 
  select(strap_number, results_tidy, results_glance) |> 
  unnest(results_tidy, results_glance)
```

```{r}
results_logb1b2 = boot_results |> 
  select(strap_number, term, estimate, std.error) |> 
  filter(term %in% c("tmin","prcp")) |> 
  pivot_wider(
     names_from = term,
     values_from = c(estimate, std.error)
  ) |> 
  mutate(
    logb1b2 = log(estimate_tmin * estimate_prcp)
  ) 

results_rsq = boot_results |> 
  select(strap_number, r.squared) |> 
  distinct(.keep_all = TRUE)
```

Looking at log(b1*b2)
```{r}
results_logb1b2 |> ggplot(aes(x = logb1b2)) + geom_boxplot() +
  labs(title = "Distribution of log(b1*b2)")
```

The median of log(b1*b2) is around -6, the data is left-skewed with many outliers at the lower end. 

One thing to note here is that there are a lot of negative estimate values for `prcp`, so many log(b1\*b2) is `NaN`, and these are excluded from plotting or calculation of quantiles. Only `r sum(!is.na(results_logb1b2$logb1b2))` rows out of `r nrow(results_logb1b2)` rows have valid log(b1\*b2) values. 

The 95% CI for log(b1*b2):
```{r}
results_logb1b2 |> summarise(
  lowerCI = quantile(logb1b2, 0.025, na.rm = TRUE),
  upperCI = quantile(logb1b2, 0.975, na.rm = TRUE)) |> 
  knitr::kable()
```

```{r}
results_rsq |> ggplot(aes(x = r.squared)) + geom_boxplot() + 
  labs(title = "Distribution of R squared")
```

The estimated R squared value is close to 1, meaning there is a high percentage in the variance of `tmax` that can be explained by the relationship with the predictors `tmin` and `prcp`. 

The 95% CI for r-squared:
```{r}
results_rsq |> summarise(
  lowerCI = quantile(r.squared, 0.025, na.rm = TRUE),
  upperCI = quantile(r.squared, 0.975, na.rm = TRUE)) |> 
  knitr::kable()
```

The 95% CI is relatively small and close to 1, meaning overall the model fits the data relatively well. 

## Problem 3

Reading and cleaning data

```{r}
bwt_df = read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = factor(babysex, levels = c(1, 2), labels = c("male", "female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
  mrace = factor(mrace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
  malform = factor(malform, levels = c(0, 1), labels = c("absent", "present"))
  )

print("Displaying the number of missing values in each column:")
colSums(is.na(bwt_df))
```

For data cleaning, I reorganized all categorical data into factor; there is no missing data in this set. 

To propose a model, my hypothesis is that birth weight is associated with presence of malformation `malform` and mother’s weight gain during pregnancy `wtgain`, since I hypothesize pregnant mom with greater weight gain may have babies that weigh more, and babies who don't have malformation might have higher weight. I will use MLR to fit `malform` and `wtgain` as predictor, and `bwt` as outcome. I will also include an interaction term since I assume `malform` may be an effect modifier. 

Fitting the model I proposed: 
```{r}
q3m1 = lm(bwt ~ malform + wtgain + malform*wtgain, data = bwt_df)

broom::tidy(q3m1)
```

Plotting residuals against fitted value for my model:
```{r}
bwt_df |> modelr::add_predictions(q3m1) |> 
  modelr::add_residuals(q3m1) |> 
  ggplot(aes(x = pred, y = resid)) + geom_point() +
  labs(title = "MLR of birth weight with malformations and weight gain")
```

From this plot, we can see that the residuals are very large, meaning the this is not an optimal model. 

Now comparing with the other two given models:
```{r}
q3model2 = lm(bwt ~ blength + gaweeks, data = bwt_df)
broom::tidy(q3model2)

q3model3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = bwt_df)
broom::tidy(q3model3)
```

```{r}
cv_df = 
  bwt_df |> 
  modelr::crossv_mc(n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_results = cv_df |> mutate(
  q3m1 = map(train, \(df) lm(bwt ~ malform + wtgain + malform*wtgain, data = df)),
  q3m2 = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
  q3m3 = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = df))
  )|> 
    mutate(
    rmse_q3m1 = map2_dbl(q3m1, test, \(mod, df) modelr::rmse(mod, df)),
    rmse_q3m2 = map2_dbl(q3m2, test, \(mod, df) modelr::rmse(mod, df)),
    rmse_q3m3 = map2_dbl(q3m3, test, \(mod, df) modelr::rmse(mod, df))
    )

cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model_type",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  group_by(model_type) |> 
  summarize(m_rmse = mean(rmse))
```

Plot the RMSE for all three models
```{r}
cv_results |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model_type",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x = model_type, y = rmse)) +
  geom_violin() +
  labs(title = "Comparison between my model (q3m1) and 2 given models")
```

The 2 alternative models are significantly better than the model I proposed due to lower RMSE. 
